---
title: "Prison mapping step 2: images export"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description


### 0 - Load packages

```{r message=FALSE}

library(dplyr)
library(sf)
library(tmap)
library(mapview)
library(readr)
library(fs)

```


### I - Get data

* Filtered out Juvenile and Closed facilities
```{r}

prison_bdys <- st_read("../../data/gis/prison_bdys.geojson", 
                       stringsAsFactors = FALSE, quiet = TRUE) %>% 
  filter(STATUS != "CLOSED", SECURELVL != "JUVENILE") %>% 
  select(FID:COUNTYFIPS, SECURELVL:CAPACITY) %>% 
  st_transform(crs = 2163)

st_crs(prison_bdys)
st_bbox(prison_bdys)

prison_ctrs_groups <- st_read("exports/prison_ctrs_groups.geojson", 
                              stringsAsFactors = FALSE, quiet = TRUE) %>% 
  st_transform(crs = 2163)

st_crs(prison_ctrs_groups)
st_bbox(prison_ctrs_groups)

```

### II - Looping for map images

* Doing these state by state (by state code), could implement differently but there are a lot of images to generate so taking time to examine more manageable subsets of them

```{r}

# the 50 states
sort(state.abb)

cat("also DC and these US territories: \n")
prison_bdys %>% 
  filter(!STATE %in% state.abb) %>% .$STATE %>% 
  unique() %>% sort()

```


* Manually pick a state by modifying the group_IDs code (here STATE == "VT", small one w/relatively few facilities to speed up this example)

* The loop utilizes the tmap package to make the map and convert it to leaflet, and the mapview package to save a webshot of it (otherwise the basemap will not show)

* labeling formatting options seem to be somewhat limited, just adding the facility ID but would like to position it more flexibly and also add a shadow/mask to help it show up against all backgrounds

* *(Experimenting with expanding the bounding box, change `tm_view(bbox = bdys_bbox2)` to use `bdys_bbox` to map just the boundary feature(s) extent)*

```{r}

stime <- Sys.time()

group_IDs <- prison_ctrs_groups %>% 
  filter(STATE == "VT") %>% 
  .$group %>% unique()


for (i in 1:length(group_IDs)) {
  prisonpoints <- prison_ctrs_groups %>% filter(group == group_IDs[i])
  prison_ids <- prisonpoints$FACILITYID
  prisonbdys <- prison_bdys %>% filter(FACILITYID %in% prison_ids)
  prison_state <- unique(prisonbdys$STATE)
  title_text <- paste(prison_state, paste0(prison_ids, collapse = " "))
  bdys_bbox <- st_bbox(prisonbdys)
  bdys_bbox2 <- st_bbox(c(bdys_bbox[1] + -100, bdys_bbox[2] + -100, 
                          bdys_bbox[3] + 100, bdys_bbox[4] + 100), 
                       crs = st_crs(2163))
  
  map <- 
    tm_view(bbox = bdys_bbox2) +
    tm_basemap(leaflet::providers$Esri.WorldImagery) + 
    tm_shape(prisonpoints) + 
    tm_dots() +
    tm_shape(prisonbdys) + 
    tm_polygons(alpha = 0, border.col = "red", lwd = 2) + 
    tm_text("FACILITYID", remove.overlap = FALSE, 
            size = 1.5, col = "black") +     
    tm_layout(title = title_text) +
    tm_scale_bar(position=c("left", "bottom"))
    lf <- tmap_leaflet(map)
    dir_create(paste0("exports/"), prison_state)
    mapshot(lf, 
            file = paste0(paste0("exports/", prison_state, "/"), 
                          prison_state, "_", group_IDs[i], ".png"),
            remove_controls = c("zoomControl", "layersControl", "homeButton"),
            zoom = 0.8)

}


prison_ctrs_groups %>% 
  filter(group %in% group_IDs) %>% 
  st_drop_geometry() %>% 
  mutate(image_name = paste0(STATE, "_", group, ".png"),
         image_path = paste0("exports/", STATE, "/", image_name)) %>% 
  write_csv(., paste0("exports/", prison_state, "/", "prisons_data_imagelinks.csv"))

etime <- Sys.time()

etime - stime

```

* Show an image

```{r message=FALSE, fig.width=8, fig.height=8}

read_csv(paste0("exports/", prison_state, "/", "prisons_data_imagelinks.csv")) %>% 
  slice(nrow(.)) %>% .$image_path %>% imager::load.image() %>% 
  plot(axes = FALSE)

```


<br>

