---
title: "Prison mapping step 1: get points"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Description

Get polygon center points, calculate distances, and get group center points.

### 0 - Load packages

```{r message=FALSE}

library(sf)
library(nngeo)
library(dplyr)
library(readr)

```

### I - Get data

* Filtered out Juvenile and Closed facilities
```{r}

prison_bdys <- st_read("../../data/gis/prison_bdys.geojson", 
                       stringsAsFactors = FALSE, quiet = TRUE) %>% 
  filter(STATUS != "CLOSED", SECURELVL != "JUVENILE") %>% 
  select(FID:COUNTYFIPS, SECURELVL:CAPACITY)

prison_bdys

```

* Some stats

```{r}

table(prison_bdys$STATUS, prison_bdys$TYPE)

table(prison_bdys$STATUS, prison_bdys$SECURELVL)

```

### II - Calculate distances, create centroids and groups

* Work in a projected coordinate system ([EPSG:2163](https://epsg.io/2163)) 

```{r warning=FALSE}

prison_ctrs <- prison_bdys %>% 
  st_transform(crs = 2163) %>% 
  st_centroid()

st_crs(prison_ctrs)

```

* Visualize extent
```{r}

plot(prison_ctrs["STATE"], key.pos = NULL, 
     main = "Prison Facilities (projected CRS)")

```


1. Calculate nearest distances, add to prison_ctrs data frame 

* *(For potential later use: e.g., identifying relatively isolated facilities > some cutoff distance)*

```{r}

nearest_1 <- st_nn(prison_ctrs, prison_ctrs, k = 2, returnDist = TRUE)

nearest_1_id <- nearest_1[["nn"]]
nearest_1_id <- do.call(rbind, nearest_1_id)[,2]
nearest_1_facid <- prison_ctrs[nearest_1_id, ]$FACILITYID
nearest_1_facid <- data.frame(nearest_1_id, nearest_1_facid, stringsAsFactors = FALSE)

nearest_1_dist <- nearest_1[["dist"]][,2]
nearest_1_dat <- cbind(nearest_1_facid, nearest_1_dist)

prison_ctrs <- st_sf(data.frame(prison_ctrs, nearest_1_dat))

```


2. Make groups based on hierarchical clustering, for aggregating near points to get a centroid (reduces the number of images since some facilities are near/next to each other)

* Note there doesn't seem to be a grouping variable for prison facilities even though some are clearly part of the same complex

* Adapted from: https://stackoverflow.com/questions/26540831/how-to-cluster-points-and-plot

* Choosing 1 km, so h = 1000

```{r}

prison_ctrs_coords <- prison_ctrs %>% st_coordinates()

dend <- hclust(dist(prison_ctrs_coords), method = "complete")

groups <- cutree(dend, h = 1000)

# number of facilities
length(groups)

# number of images (eventually)
max(groups)

prison_ctrs_coordgroups <- 
  tibble(FACILITYID = prison_ctrs$FACILITYID, 
         X = prison_ctrs_coords[,1], 
         Y = prison_ctrs_coords[,2], 
         group = groups) %>% 
  arrange(group, FACILITYID)

head(prison_ctrs_coordgroups)

```

* Loop to get bounding box center points for each group

```{r}

max_group <- max(prison_ctrs_coordgroups$group)

group_cp <- list()

for(i in 1:max_group) {
  
  pts <- prison_ctrs_coordgroups %>% filter(group == i)
  pts_sf <- st_as_sf(pts, coords = c("X","Y"))
  pts_ext <- st_bbox(pts_sf)
  pts_ext_cp <- c(mean(pts_ext[c(1, 3)]), mean(pts_ext[c(2, 4)]))
  group_cp[[i]] <- cbind(group = i, cp_X = pts_ext_cp[1], cp_Y = pts_ext_cp[2])
  
}

group_cp_df <- do.call(rbind, group_cp) %>% as_tibble()


prison_ctrs_coordgroups <- left_join(prison_ctrs_coordgroups, 
                                           group_cp_df, by = "group")

```


* Quick test if they work

```{r}

prison_ctrs_coordgroups %>% group_by(group) %>% 
  tally() %>% arrange(-n) %>% slice(1:5)

# pick one, the center point should be in the middle of the facility centroids

plot(prison_ctrs_coordgroups[prison_ctrs_coordgroups$group == 79,]$Y, 
     prison_ctrs_coordgroups[prison_ctrs_coordgroups$group == 79,]$X, 
     xlab = "Longitude (m)", 
     ylab = "Latitude (m)")
points(prison_ctrs_coordgroups[prison_ctrs_coordgroups$group == 79,]$cp_Y, 
       prison_ctrs_coordgroups[prison_ctrs_coordgroups$group == 79,]$cp_X,
       col = "red", pch = 16)

```

* Export for further work (note that geojson crs should be 4326)

```{r}

prison_ctrs %>% st_transform(crs = 4326) %>% 
  inner_join(., prison_ctrs_coordgroups, by = "FACILITYID") %>% 
  st_write("exports/prison_ctrs_groups.geojson", delete_dsn = TRUE)

```

#### Next step: generating images with a basemap

<br>

